---
title: 'Intervals and Predictive Distributions'
author: "Professor Rodrigo Targino"
output:
  xaringan::moon_reader:
    css: ['default', "xaringan_style.css"]
---

```{r, echo=FALSE, warning=FALSE}
library(knitr)
library(ggplot2)
library(kableExtra)
suppressPackageStartupMessages(library(tidyverse))
opts_chunk$set(echo=FALSE, fig.align='center', out.width="60%") 
knitr::opts_chunk$set(dpi=300, fig.width=7)
#options("kableExtra.html.bsTable" = T)
ig <- function(file) {knitr::include_graphics(file)}
source("../scripts/make_figs.R")
source("../scripts/color_defs.R")

```

### Announcements

- Reading: Chapters 8.1 (intervals) and 8.3 (posterior prediction) Bayes Rules

- Homework 2 due May 5, at 11:59pm

---

### Reminder: Frequentist confidence interval

- Frequentist interval: $Pr(l(Y) < \theta < u(Y)\mid \theta) = 0.95$

    +  Probability that the interval will cover the true value _before_ the data are observed. 
    
    + Interval is random since $Y$ is random

---

### Reminder: Frequentist confidence interval

```{r, out.width='60%'}
par(cex=1.5)
prob <- 0.6
nintervals <- 50
sz <- 20
counts <- rbinom(n=nintervals, size=sz, prob=prob)
phat <- counts / sz

lower <- phat - 1.96 * sqrt(phat * (1-phat) / sz)
upper <- phat + 1.96 * sqrt(phat * (1-phat) / sz)

plot(0, 0, ylim=c(-1, nintervals + 1), xlim=c(0, 1), cex=0, xlab=expression(theta), ylab="")
segments(x0=lower, x1=upper, y0=1:nintervals, y1=1:nintervals, col=ifelse(upper < prob | lower > prob, cols2[1], "grey"), lwd=3)
abline(v=prob, col=cols2[2])

```

.center[We expect 0.05 x 50 = 2.5 will _not_ cover the true parameter 0.6]

---

### Posterior Credible Intervals

- Frequentist interval: $Pr(l(Y) < \theta < u(Y)\mid \theta) = 0.95$

    +  Probability that the interval will cover the true value _before_ the data are observed. 
    
    + Interval is random since $Y$ is random

--
    
- **Bayesian Interval**: $Pr(l(y) < \theta < u(y) \mid Y=y) = 0.95$
    
    +  Information about the the true value of $\theta$ _after_ observeing $Y = y$.
    
    + $\theta$ is random (because we include a prior), $y$ is observed so interval is non-random.  

---

### Posterior Credible Intervals (Quantile-based)

- The easiest way to obtain a confidence interval is to use the quantiles of the posterior distribution.

If we want $100 \times (1 - \alpha)%$ interval, we find numbers $\theta_{\alpha / 2}$ and $\theta_{1-\alpha/2}$ such that:

1. $p(\theta \lt \theta_{\alpha/2} \mid Y=y) = \alpha/2$

2. $p(\theta \gt \theta_{1-\alpha/2} \mid Y=y) = \alpha/2$

$$p(\theta \in [\theta_{\alpha/2}, \theta_{1-\alpha/2}] \mid Y=y) = 1 - \alpha$$

- Use quantile functions in R, e.g. `qbeta`, `qpois`, `qnorm` etc.  

---

### Example: interval for shooting skill in basketball

- The posterior distribution for Covington's shooting percentage is a $$\text{Beta}(49 + 478, 50 + 873) = \text{Beta}(528, 924)$$

- For a 95% _credible_ interval, $\alpha = 0.05$
   + Lower endpoint: `qbeta(0.025, 528, 924)`
   + Upper endpoint: `qbeta(0.975, 528, 924)`
   +  $[\theta_{\alpha/2}, \theta_{1-\alpha/2}]$ =  [`r round(qbeta(0.025, 528, 924), 2)`, `r round(qbeta(0.975, 528, 924), 2)`]

--

- Compared to frequentist _confidence_ interval without prior information: [`r round(0.49 - 1.96*sqrt((0.49 * (1-0.49))/100), 2)`, `r round(0.49 + 1.96*sqrt((0.49 * (1-0.49))/101), 2)`]

- End-of-season percentage was $0.37$

- Credible intervals and confidence intervals have different meanings!

---

### Highest Posterior Density (HPD) region

**Definition: (HPD region)** A $100  \times (1 − \alpha)%$ HPD region consists of a subset of the parameter space, $R(y) \in \Theta$  such that

1. $\text{Pr}(\theta \in R(y) | Y = y ) = 1 - \alpha$

   - The probability that $\theta$ is in the HPD region is $1 - \alpha$

--

2. $\text{If }\theta_{a} \in R(y) , \text{ and } \theta_{b} \notin R(y) , \text{ then } p\left( \theta_{ a } | Y = y \right) > p \left( \theta_{ b } | Y = y \right)$

   - All points in an HPD region have a higher posterior density than points out- side the region.

The HPD region can be discontinuous (hence "region")

---

### Highest Posterior Density (HPD) region

1. $p(\theta \in s(y) \mid Y = y) =1- \alpha$

2. If $\theta_a \in s(y)$, and $\theta_b \not \in s(y)$, then $p(\theta_a \mid Y=y) \gt p(\theta_b \mid Y = y)$.
    
    - All points in an HPD region have a higher posterior density than points out- side the region.

<br>

.center[The  HPD region is the _smallest_ region with probability `\\((1 - \alpha)\\)`% ]

---

### Highest Posterior Density (HPD) region

```{r, out.width='60%'}
library("HDInterval")
data <- c(rgamma(1e5, 3, 3), rgamma(1e5, 20, 2))

compute_hpd = function(alpha, data){
  dens <- density(data)
  plot(dens, lwd=2, typ="l", xlab=expression(theta), ylab= "Posterior density", 
       main = bquote("HPD with " ~ alpha == .(alpha)))
  
  hdiD2 <- hdi(dens, allowSplit=TRUE, credMass = 1-alpha)
  ht <- attr(hdiD2, "height")
  
  segments(hdiD2[, 1], ht, hdiD2[, 2], ht, lwd=3, col='red')
}

alpha = 0.75

compute_hpd(alpha, data)
```

---

### Highest Posterior Density (HPD) region

```{r, out.width='60%'}
alpha = 0.6

compute_hpd(alpha, data)
```

---

### Highest Posterior Density (HPD) region

```{r, out.width='60%'}
alpha = 0.5

compute_hpd(alpha, data)
```

---

### Highest Posterior Density (HPD) region

```{r, out.width='60%'}
alpha = 0.25

compute_hpd(alpha, data)
```

---

### Highest Posterior Density (HPD) region

```{r, out.width='60%'}
alpha = 0.1

compute_hpd(alpha, data)
```

---

### Highest Posterior Density (HPD) region

```{r, out.width='60%'}
alpha = 0.05

compute_hpd(alpha, data)
```

---

### Calibration: Frequentist Behavior of Bayesian Intervals

<br>

- A credible interval is calibrated if it has the right frequentist coverage

- Bayesian credible intervals usually won't have correct coverage

- If our prior was well-calibrated and the sampling model was correct, we'd have well-calibrated credible intervals

- Specifying _nearly_ calibrated prior distributions is hard!

---

### Calibration of MLB predictions

<br>
<br>

```{r, out.width="85%"}
ig("images/538_calibration1.png")
```
.center[source: https://projects.fivethirtyeight.com/checking-our-work/]

---

### Calibration of MLB predictions

```{r, out.width="60%"}
ig("images/538_calibration1_2.png")
```
.center[source: https://projects.fivethirtyeight.com/checking-our-work/]

---
### Calibration of political predictions

<br>

```{r, out.width="85%"}
ig("images/538_calibration2.png")
```
.center[source: https://fivethirtyeight.com/features/when-we-say-70-percent-it-really-means-70-percent/]

---

### The age guessing game*

<br>

```{r, out.width='35%'}
ig("images/age_guess.jpg")
# Guess Prof Franks' age to the nearest (months)

###Suppose I told you I'm between 10 and 50.
### Suppose I told you I'm between 30 and 33 y/o? (360 and 396


```

.center[ *Bayesian edition]

---

### Interval Trivia

- I'm going to ask you ten questions about random facts

- For each, write down a 50% credible interval for _your_ belief about the answer
  
- Goal:
  + Be well calibrated. 50% of your intervals should contain the true answer.

---

### Percentage in California
```{r, out.width="70%"}
ig("images/california.png")
```

.center[What fraction of the US population is living in California?]

---

### Olympic Swimming Pool

```{r, out.width="80%"}
ig("images/pool.jpeg")
```

.center[How many gallons of water are there in an olympic swimming pool?]

---

### Household Income

<br>

```{r, out.width="80%"}
ig("images/householdincome.jpg")
```

<br>

.center[What is the median household income in the US (dollars)?]

---

### Grooves on a quarter

<br>

```{r, out.width="80%"}
ig("images/quarter.jpeg")
```

.center[How may grooves are there on the edge of quarter?]


---

### Gold in Fort Knox

<br>

```{r, out.width="70%"}
ig("images/gold.jpg")
```

.center[How many pounds of gold are there Fort Knox?]

---

### Population of Australia

```{r, out.width="55%"}
ig("images/australia.png")
```

.center[What is the population of Australia (in millions)?]

---

### Tallest tsunami wave ever recorded

<br>

```{r}
ig("images/tsunami.jpg")
```

.center[How tall was the tallest tsunami wave ever recorded (in feet)?]

---

### Disney world

<br>

```{r}
ig("images/disney.jpg")
```

.center[In what year did Disneyland Park open to the public?]

---

### Jupiter

<br>

```{r}
ig("images/jupiter.jpg")
```

.center[How many times larger in volume is Jupiter than Earth?]

---

### Netflix

<br>

```{r}
ig("images/netflix.png")
```

.center[In what year was Netflix founded?]

---

###  All questions

```{r}
questions <- c("What fraction of the US population is living in California?",
  "How many gallons of water are there in an olympic swimming pool?",
  "What is the median household income in the US (dollars)?",
  "How many grooves are there on the edge of quarter?",
  "How many pounds of gold are there in Fort Knox?",
  "What is the population of Australia (in millions)?",
  "How tall was the tallest tsunami wave ever recorded (in feet)? ", 
  "In what year did Disneyland open to the public?",
  "How many times larger in volume, is Jupiter than Earth?",
  "In what year was Netflix founded?")
lower <- upper <- rep("______", length(questions))

kable(tibble(N=paste0("Q",1:10), Question=questions, lower=lower, upper=upper)) %>% kable_styling(font_size = 16)

```

---

###  Answers
```{r}
Answers <- c("12%", "660,000", "$70,784", "119", "9,206,250", "25.89 mil", "1700 feet", "1955", "1,321.33", "1997")

kable(tibble(N=paste0("Q",1:10), Question=questions, Answers=Answers)) %>% kable_styling(font_size = 16)
```

---

###  References
```{r}
references <- c("https://www.census.gov/library/stories/state-by-state/california-population-change-between-census-decade.html",
  "https://healingpicks.com/how-many-gallons-are-in-an-olympic-size-swimming-pool/",
  "https://www.census.gov/library/visualizations/2022/comm/median-household-income.html",
  "https://www.icpl.org/articles/why-does-quarter-have-119-ridges",
  "https://www.usmint.gov/about/mint-tours-facilities/fort-knox",
  "https://www.abs.gov.au/statistics/people/population",
  "https://www.pbs.org/wgbh/nova/tsunami/once-nf.html", 
  "https://www.history.com/this-day-in-history/disneyland-opens",
  "https://www.space.com/18392-how-big-is-jupiter.html",
  "https://en.wikipedia.org/wiki/Netflix")

kable(tibble(N=paste0("Q",1:10), References=references)) %>% kable_styling(font_size = 16)
```

---
### Calibrated probability intervals

- Calibration is important but only part of the story!

- Want well calibrated but _small_ intervals (big intervals tell us nothing)

- How you could have gotten a perfect score on the quiz:

  + For 5 of the answers select [-1 Trillion, +1 Trill] (ensures it will cover)
  + For the other 5 answers select[-0.01, + 0.01] (ensures it won't cover)

--

- Domain expertise helps us develop smaller prior distributions (calibration?)
    
  + Usually at the cost of calibration
  + My experience: people tend to be overconfident
  + Alternatives to domain expertise?

---

### Subjective Bayesianism

- So far we have focused on defining priors using domain expertise

- "Subjective" Bayes

    + Essentially what we have discussed so far

    + Priors usually represent subjective judgements can't always be rigorously justified 
    
- Alternative: "objective" Bayes

---

### Objective Bayesianism

- Is there a way to define "objective" prior distributions?
  
  + Good default prior distributions for some problems?
  
  + "Non-informative" prior distributions?

- Also called "reference" or "default" priors

--

- Can we find prior distributions that lead to (approximately) correct frequentist calibration?

- Can we find prior distributions which minimize the amount of information contained in the distribution?

  + Principle of maximum entropy (MAXENT).
     
---



### Difficulties with non-informative priors

.center[Uniform distribution for p]

```{r, echo=TRUE, fig.width=7, fig.height=5}
p <- runif(100)
tibble(p=p) %>% ggplot() + 
  geom_histogram(aes(x=p), bins=30) +
  theme_bw(base_size=24)
```

---

### Difficulties with non-informative priors

.center[Implied distribution for odds = p/(1-p)]

```{r, echo=TRUE, warning=FALSE, message = FALSE, fig.width=7, fig.height=5}
odds <- p/(1-p)
tibble(odds=odds) %>% ggplot() + 
  geom_histogram(aes(x=odds)) + 
  theme_bw(base_size=24)
```
---

## Improper prior distributions
- For the Beta distribution we chose a uniform prior $-$ e.g. $p(\theta) \propto   \text{const}$.  This was ok because

  + $\int_0^1 p(\theta) d\theta = \text{const} < \infty$

  + We say this prior distribution is _proper_ because it is integrable
    
- For the Poisson distribution, try the same thing: $p(\lambda) \propto \text{const}$
    + $\int_{0}^{\infty} p(\lambda) d\lambda = \infty$
    + In this case we say $p(\lambda)$ is an _improper_ prior

---

## Improper prior distributions
- In the Poisson case, let us pretend the prior for $\lambda$ is proper...
$$p(\lambda \mid y) \propto \phantom{.......................................................................................}$$
---

## Improper prior distributions

- Sometimes there is an absence of precise prior information

- The prior distribution does not have to be proper but the posterior does!

    + A proper distribution is one with an integrable density

    + If you use an improper prior distribution, you need to check that the posterior distribution is also proper


---

layout: false
class: middle, center
background-image: none;

# Posterior Predictive Distributions



---

### Posterior predictive distribution

- An important feature of Bayesian inference is the existence of a predictive distribution for new observations.
  
    + Let $\tilde y$ be a new (unseen) observation, and $y_1, ... y_n$ the observed data.
  
    + The posterior predictive distribution is $p(\tilde y \mid y_1, ... y_n)$

--

- The predictive distribution does not depend on unknown parameters

- The predictive distribution only depends on observed data

- Asks: what is the probability distribution for new data given observations of old data?

---

### Another Basketball Example

- I take free throw shots and make 1 out of 2.  How many do you think I will make if I take 10 more?

- If my true "skill" was 50%, then $\tilde Y \sim \text{Bin}(10, 0.50)$ 

- Is this the correct way to calculate the predictive distribution?


---

### Posterior Prediction

If you know $\theta$, then we know the distribution over future attempts:
$$\tilde Y \sim \text{Bin}(10, \theta)$$

```{r, warning=FALSE, out.width="80%", fig.width=10}

library("RColorBrewer")

nobs <- 2
yobs <- 1

npred <- 10

alpha <- 1
beta <- 3

theta_grid <- seq(0.1, 0.9, by=0.1)
weight <- dbeta(theta_grid, yobs + alpha, nobs - yobs + beta)
weight <- weight / sum(weight)
tib <- tibble(x=rep((0:npred)/npred, length(theta_grid)), 
             density=dbinom(npred*x, npred, rep(theta_grid, each=npred+1)), 
            theta=factor(rep(theta_grid, each=npred + 1)),
            weight = rep(weight, each=npred+1))
        
tib %>%
  ggplot() + 
  geom_line(aes(x=10*x, y=density, colour=theta), size=1.1) + 
  xlim(c(0, 10)) + theme_bw(base_size=24) + xlab("Makes") + ylab("") + scale_colour_manual(values = rev(brewer.pal(9,"RdBu"))) +
  theme(legend.position="bottom")
```

---

### Posterior Prediction

- We already observed 1 make out of 2 tries.

- Assume a Beta(1, 3) prior distribution 

  + e.g. a priori you think I'm more likely to make 25% of my shots

- Then $p(\theta \mid Y=1, n=2)$ is a $\text{Beta}(2, 4)$

- Intuition: weight $\tilde Y \sim \text{Bin}(10, \theta)$ by $p(\theta \mid Y=1, n=2)$

---

### Posterior Prediction

.center[If I take 10 more shots how many will I make?]

```{r, dependson=-1, out.width="80%", fig.width=10}
  tib %>%
  ggplot() + 
  geom_line(aes(x=10*x, y=weight*density, colour=theta), size=1.1) + 
  xlim(c(0, 10)) + theme_bw(base_size=24) + xlab("Makes") + ylab("") + 
  scale_colour_manual(values = rev(brewer.pal(9,"RdBu"))) + 
  theme(legend.position="bottom")
```

---

### Posterior Predictive Distribution

$$p(\theta) = \text{Beta}(1, 3), p(\theta \mid y) = \text{Beta}(2, 4)$$

```{r, dependson=-3, warning=FALSE, out.width="50%"}
library("RColorBrewer")
nobs <- 2
yobs <- 1

npred <- 10

# alpha <- 2
# beta <- 2

###alpha <- 1
###beta <- 1

dbetabinom <- function(x) {exp(lchoose(npred, x) + lbeta(x + alpha + yobs, npred - x + beta + nobs-yobs) - lbeta(alpha + yobs, beta + nobs-yobs))}

tibble(x=(0:npred)/npred, Predictive = dbetabinom(npred*x),
                              `Bin(10, 0.5)` = dbinom(npred*x, npred, yobs/nobs)) %>%
  gather(key=Type, value=Height, -x) %>%
  ggplot() + 
  geom_line(aes(x=10*x, y=Height, colour=Type), size=1.1) + 
  xlim(c(0, 10)) + theme_bw(base_size=24) + xlab("Makes") + ylab("") +
  theme(legend.position = c(0.8, 0.8),
        legend.background = element_rect(fill = alpha("white", 0)))

```
The predictive density, $p(\tilde y \mid y)$, answers the question "if I take 10 more shots how many will I make, given that I already made 1 of 2".

---

### The posterior predictive distribution

$$\begin{aligned}
p(\tilde y \mid y_1, ... y_n) &= \int p(\tilde y, \theta \mid y_1, ... y_n) d\theta\\
&=\int p(\tilde y \mid \theta) p(\theta  \mid y_1, ... y_n) d\theta
\end{aligned}$$

- The posterior predictive distribution describes our uncertainty about a new observation after seeing $n$ observations

- It incorporates uncertainty due to the sampling in a model $p(\tilde y \mid \theta)$ _and_ our posterior uncertainty about the data generating parameter, $p(\theta \mid y_1, ... y_n)$


---
### Posterior Predictive Density

<br>

```{r, out.width="80%"}
ig("images/central_dogma_inference.png")
```


---

### The prior predictive distribution

$$\begin{aligned} p(\tilde y) &= \int p(\tilde y, \theta) d\theta\\ &=\int p(\tilde y \mid \theta) p(\theta) d\theta \end{aligned}$$

- The prior predictive distribution describes our uncertainty about a new observation before seeing data

--

- It incorporates uncertainty due to the sampling in a model $p(\tilde y \mid \theta)$ _and_ our prior uncertainty about the data generating parameter, $p(\theta)$

---

### Homework 1

- $\lambda \sim \text{Gamma}(\alpha, \beta)$

- $\tilde Y \sim \text{Pois}(\lambda)$

-  $p(\tilde y) = \int p(\tilde y \mid \lambda) p(\lambda) d\lambda$ is a prior predictive distribution!

- "A Gamma-Poisson mixture is a Negative-Binomial Distribution"

---

### Homework 1

$$\begin{aligned}
p(\tilde y) &= \int p(\tilde y \mid \lambda) p(\lambda) d\lambda\\
&= \int (\frac{\lambda^{\tilde y}}{y!}e^{-\lambda})(\frac{\beta^\alpha}{\Gamma(\alpha)}\lambda^{(\alpha-1)}e^{-\beta\lambda}) d\lambda\\
&= \frac{\beta^\alpha}{\Gamma(\alpha)y!}\int (\lambda^{(\alpha+y-1)}e^{-(\beta+1)\lambda}) d\lambda\\
\end{aligned}$$

$\int (\lambda^{(\alpha+y-1)}e^{-(\beta+1)\lambda}) d\lambda$ looks like an unormalized $\text{Gamma}(\alpha+y, \beta + 1)$


---

### Summary


- Bayesian credible intervals

     + Posterior probability that the value falls in the interval
     + Still strive for well-calibrated intervals (in the frequentist sense)

- Non-informative prior distributions

- Posterior predictive distributions
     + Estimated distribution for new data  our uncertainty about the parameters
     
```{css, echo=FALSE}
@media print {
  .has-continuation {
    display: block;
  }
}
```